\section{Extensions and Perspective}
We note that our techniques for $\mathbf{S}$-Robust Non-uniform Ultimate Positivity hinge on the First Order Theory of the Reals. Observe that this was rather agnostic to the exact shape of the neighbourhood: we can easily extend the same techniques to arbitrary \textit{semi-algebraic} neighbourhoods. The uniform variants, on the other hand, can easily be seen to reduce to Positivity and Ultimate Positivity when the neighbourhoods are polyhedra. Let $\mathcal{P}$ be a polyhedron that contains the origin, and let $\mathbf{d_1}, \dots, \mathbf{d_k}$ be its corners. Our critical inequality \ref{eq:crux} in this case is:
\begin{equation}
\seq{\mathbf{p}, \mathbf{q_n}} \ge \max_{\mathbf{d} \in \mathcal{P}} \seq{\mathbf{d}, \mathbf{q_n}}
\end{equation}
The inner product on the right is always maximised at a \textit{corner}. Thus, it is equivalent to ask whether for all $j \in \{1, \dots, k\}$,
\begin{equation}
\seq{\mathbf{p} - \mathbf{d_j}, \mathbf{q_n}} \ge 0
\end{equation}
These are precisely $k$ many instances of (Ultimate) Positivity, which are decidable up to order five, and number-theoretically hard at order six \cite{joeljames3}.

As outlined at the outset, we contributed towards a sharp and comprehensive picture of what is \textit{decidable} about Robust Positivity Problems for real algebraic Linear Recurrence Sequences. We find it remarkable that number-theoretic analyses involving Diophantine approximation, which usually show up in the context of hardness, also play a significant role in our \textit{decidability} proofs! However, a rather conspicuous gap in our picture is the status of $\mathbf{S}$-Robust Non-uniform Ultimate Positivity at order $5$: this seems to require even more delicate analysis. 

An obvious, but possibly tedious future direction would be to tie up the book-keeping loose ends, and meticulously account for the complexity of our techniques. We chose to work with algebraic numbers; in settings involving rational numbers where scaling to integers and accessing an $\mathsf{PosSLP}$ oracle is viable, the complexity usually lies in $\mathsf{PSPACE}$. However, this might blow up significantly in the absence of efficient positivity testing for a different class of arithmetic circuit.

At a higher level, we note that we chose our norm to be based on the standard matrix inner product. It is interesting to investigate what kinds of decidability and hardness results hold for neighbourhoods specified using different norms. Perhaps, results could be universal across a wider class of norms, and there could be a profound underlying linear-algebraic reason whose discovery would be mathematically significant.

In the grand Formal Methods scheme, the study of Hyper-properties \cite{hyperproperties} is an exciting natural way robustness problems for Linear Dynamical Systems could fit in. Hyper-logics reason about sets of traces of an infinite time system, rather than a single trace. They gained importance as a means to verify security in view of attacks like Meltdown and Spectre. A quintessential hyper-property, for instance, would specify a reasonable notion of \textit{indistinguishability} of traces. In that regard, our notions of $\mathbf{S}$-Robust Positivity and $\mathbf{S}$-Robust Uniform Ultimate Positivity bear striking resemblance. Exploring deeper connections is a fascinating future research avenue.
